---
title: "Reproducible Research: Peer Assessment 1"
author: "Michael O'Connor"
date: "December 7, 2014"
output:
  html_document:
    keep_md: true
---

## Loading and preprocessing the data

First we'll load the `activity.csv` data.  I downloaded it to my
`~/Downloads` directory, so we'll load it from there and then convert
the `date` column to dates with `as.Date`.

```{r}
data <- read.csv("~/Downloads/activity.csv")
data$date <- as.Date(data$date)
```

## What is mean total number of steps taken per day?

We'd like to look at how many steps were taken over whole days,
so let's first compute that (ignoring `NA`s).

```{r}
stepsPerDay <- tapply(data$steps,data$date,sum,na.rm=TRUE)
```

Now let's see a histogram of total steps per day:

```{r}
hist(stepsPerDay,xlab="Steps Per Day",main="Histogram of Steps Per Day")
```

The mean and median steps per day are as follows (again, ignoring
`NA`s):

```{r}
list(mean=mean(stepsPerDay),median=median(stepsPerDay))
```

## What is the average daily activity pattern?

For the average daily activity pattern, we want to average the number
of steps over all days for each 5-minute interval, then graph a
timeseries of that.

The only complication is that for the plot we want now want to convert
the intervals (which are integers in the format `%H%M`) to regular times.

```{r}
averagedSteps <- tapply(data$steps,data$interval,mean,na.rm=TRUE)

times <- strptime(sprintf("%04d",as.integer(dimnames(averagedSteps)[[1]])),format="%H%M")

plot(times,averagedSteps,type="l",xlab="",ylab="Steps",main="Steps Averaged Over 5-Minute Intervals",col="blue",lwd=3,tck=1)
```


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
