---
title: "Reproducible Research: Peer Assessment 1"
author: "Michael O'Connor"
date: "December 7, 2014"
output:
  html_document:
    keep_md: true
---

## Loading and preprocessing the data

First we'll load the `activity.csv` data.  I downloaded it to my
`~/Downloads` directory, so we'll load it from there and also convert
the `date` and `interval` columns.

The `date` column can just be converted with `as.Date`.  The
`interval` column contains times of day (without dates) in the format
`%H%M`, so it'll be just a little more work to convert that correctly. 

```{r}
data <- read.csv("~/Downloads/activity.csv")
data$date <- as.Date(data$date)
data$interval <- strptime(sprintf("%04d",data$interval),format="%H%M")
```

## What is mean total number of steps taken per day?

Let's see a histogram of total steps per day (ignoring `NA`s):

```{r}
with(data,hist(steps,col="red"))
```

The mean and median steps per day are as follows (again, ignoring
`NA`s):

```{r}
list(mean=mean(data$steps,na.rm=TRUE),median=median(data$steps,na.rm=TRUE))
```


## What is the average daily activity pattern?

For the average daily activity pattern, we want to average the number
of steps over all days for each 5-minute interval, then graph a
timeseries of that.

```{r}
averagedSteps <- tapply(data$steps,data$interval,mean,na.rm=TRUE)
plot(averagedSteps)
```


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
